---
title: "Data"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, echo=FALSE}
repo.url <- 'https://github.com/liangyy/phenotype-epistasis/raw/master/'
```
 
# Intermediate GWAS

```{r, echo=FALSE}
library(pander)
panderOptions('table.split.table', Inf)
data.inter <- data.frame(id = c(), phenotype = c(), num.of.traits = c(), sample.size = c(), num.of.variants = c(), heritability.mean = c(), ancestry = c())
genDataLine <- function(id, phenotype, num.of.traits, sample.size, num.of.variants, ancestry, heritability.mean = NA) {
  data.inter <- data.frame(id = id, phenotype = phenotype, num.of.traits = num.of.traits, sample.size = sample.size, num.of.variants = num.of.variants, heritability.mean = heritability.mean, ancestry = ancestry)
  return(data.inter)
}
id.counter <- 1
```

## Blood cell phenotype

```{r, echo=FALSE}
data.inter.blood <- genDataLine(id = c(id.counter), phenotype = c('blood cell traits'), num.of.traits = c(36), sample.size = c(173480), num.of.variants = c(29.5e6), heritability.mean = c(NA), ancestry = c('european'))
id.counter <- id.counter + 1
```

* [Data](ftp://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/AstleWJ_27863252_GCST004602)
* [Paper](http://www.cell.com/cell/fulltext/S0092-8674(16)31463-5)
* Meta-data

```{r, echo=FALSE}
pander(data.inter.blood)
```


The phenotype information is as follow (Table S1)
```{r, echo=FALSE}
library(dplyr)
library(kableExtra)
library(knitr)
data <- read.csv('../external-data/10.1016.j.cell.2016.10.042-mmc1.csv')
kable(data, 'html') %>%
  kable_styling() %>%
  scroll_box(width = '1000px', height = '500px')
```

## Circulating metabolites

```{r, echo=FALSE}
data.inter.metabolite <- genDataLine(id = c(id.counter), phenotype = c('circulating metabolites'), num.of.traits = c(123), sample.size = c(24925), num.of.variants = c(39e6), heritability.mean = c(NA), ancestry = c('european'))
id.counter <- id.counter + 1
```

* [Data](http://www.computationalmedicine.fi/data#NMR_GWAS)
* [Paper](https://www.nature.com/articles/ncomms11122)
* Meta-data

```{r, echo=FALSE}
pander(data.inter.metabolite)
```


The phenotype information is as follow (Table S1)
```{r, dev='pdf', out.width = "1000px", out.height = "500px"}
knitr::include_graphics(paste0(repo.url, "external-data/ncomms11122-s1-table-s1.pdf"))
```

## Personality

```{r, echo=FALSE}
data.inter.personality <- genDataLine(id = c(id.counter), phenotype = c('personality'), num.of.traits = c(5), sample.size = c(17375), num.of.variants = c(NA), heritability.mean = c(NA), ancestry = c('european'))
id.counter <- id.counter + 1
```

* [Paper-complete](https://www.nature.com/articles/ng.3736)
* [Paper-summary statistic available](http://www.tweelingenregister.org/fileadmin/user_upload/publicaties/verslaggeving/NTR-publicaties_2012/Moor_MP_2012.pdf)
* [Data](http://www.tweelingenregister.org/GPC/)
* Meta-data

```{r, echo=FALSE}
pander(data.inter.personality)
```

The phenotype information is as follow (Table S2)
```{r, dev='pdf', out.width = "1000px", out.height = "500px"}
knitr::include_graphics(paste0(repo.url, "external-data/ng.3736-S1-table-s2.pdf"))
```

Keep in mind that the sample size is 17375 (63661 for neuroticism and extraversion) even if it is explicitly told in the readme file).